---
title: "1.1 Replica Data"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
# Explore Replica Datasets

```{r}
alameda_weekday_spend <- read_csv('data/trends-spend-merchant-v2_from-month-of-2019-01-01-to-month-of-2024-07-31--weekday_alameda')

bay_area_od_trips <- read_csv('data/trends-trip-count-od-origin-v2--mode_from-month-of-2024-02-01-to-month-of-2024-06-30--weekend_sfohc/trends-trip-count-od-origin-v2--mode_from-month-of-2024-02-01-to-month-of-2024-06-30--weekend_sfohc.csv')

bay_area_od_trips # 7.69M rows

# Get unique counties for bay area
BAY_AREA_COUNTIES_FIPS <- bay_area_od_trips$origin_geo_id %>% substr(1,5) %>% unique() # "06013" "06001" "06081" "06075" "06041"
EASTBAY_COUNTIES_FIPS <- c('06013', '06001')# "06081" "06075" "06041"

bay_area_od_trips.intra <- bay_area_od_trips %>%
  filter(substr(destination_geo_id, 1,5) %in% BAY_AREA_COUNTIES_FIPS)

bay_area_od_trips.intra # 3.5M rows

# Get Census tract geographies
CA_tracts <- sf::read_sf('data/replica-bulk-census-geographies-cal-nev/trct.geojson')

# bay_area_od_trips.intra
bay_area_od_trips.intra.grpd <- 
  bay_area_od_trips.intra %>%
  group_by(origin_geo_id, destination_geo_id) %>%
  summarize(across(ends_with('_count'), \(x) sum(x, na.rm=T)))

bay_area_tracts.ods <- 
  CA_tracts %>%
  inner_join(
    bay_area_od_trips.intra.grpd,
    by = c('id' = 'origin_geo_id')
  )

bay_area_tracts.ods %>%
  ggplot(aes(fill = biking_trip_count)) + 
  geom_sf() #+ 
  #coord_sf(xlim = c(-122.4, -122.05), ylim = c(37.7, 37.9))

  
```



Ok, so my goal is to look at tract->tract active transportation trips

Data Source: Replica

Scope:
  - 4 month period Feb-July (2024)
  - within the SF-Oakland-Hayward MSA (probably easier to jsut look at east bay)
  
  
Data Files:
  - tract geogrpahies: Replica
  - O-D pairs: Replica
  - Demographic controls: Census/IPUMS
  - Elevated infrastructure: TIGER
    - Primary Road: MTFCC = 'S1100'
    - BART lines
  - Route-generation: ??
  

Infrastructure List:
  - BART (everywhere from Hopkins north, Alcatraz south to 47th, thru West Oakland, etc) - note, the presence of the Ohlone greenway probably complicates this
  - I-80, I-880, I-570, I-980, CA-13, CA-24, I-680, 101
  - 2) CA-13, CA-123, 
  
  
Infrastructure decision points:
  - I think I want to make a tier list, and run the same analysis with multiple tiers
    - 1) interstates and elevated state highways
    - 2) BART and surface-level state highways
    - 3) ...?
    
    
It may be easy enough to just create sectors of the East Bay that are segmented (divided) by these infrastructure segements and then just use a binary on if the tracts are adjacent or 


Data Needs:
  - Distance between block groups (crow flies -- this is an opportunity for future study)
  
  
Additional Data Options:
  - Walk Score?
  -  
  
## Limit to just the East Bay
Explore how much 
```{r explore dpet}
eb_od_trips <- 
  bay_area_od_trips %>%
  # substr(geoid, 1, 5) == fips5
  filter(substr(destination_geo_id, 1, 5) %in% EASTBAY_COUNTIES_FIPS, substr(origin_geo_id, 1, 5) %in% EASTBAY_COUNTIES_FIPS)

eb_od_trips # 1.369M total O-D pairs (4 months so more like 342k)
eb_od_trips %>% distinct(origin_geo_id, destination_geo_id) # 314k

eb_od_trips %>% 
  group_by(origin_geo_id, destination_geo_id) %>%
  summarize(across(ends_with('_trip_count'), \(x) sum(x, na.rm=T))) %>%
  mutate(active_trip_count = walking_trip_count + biking_trip_count) %>%
  filter(active_trip_count > 0) # 30,728

# So only 10% of O-D pairs have any active trips between them. That's probably reasonable. Let's go back and add back in SF
```